### 2013-10-10 Publications/2013-ISBI14/03-methods.tex
Updating for second deadline

\subsection{Digital Phantom}
Based on the fiber geometries of the digital phantom 
created for the \emph{High angular resolution diffusion imaging 
reconstruction Challenge} held in ISBI 2013 
(San Francisco, US), we simulated high resolution 
(0.5mm isotropic) T1 (TE/TR=10/1500ms) and T2 
(TE/TR=90/5000ms) images, as well as a \gls*{dmri}
(1.0mm isotropic, $b$=1200, 32 evenly-distributed 
directions, 1 \textit{b0} volume) image.
The phantom includes \gls*{wm} fiber bundles, 
\gls*{csf} and \gls*{gm}. Diffusion is modeled by a 
restricted and a hindered compartment, similar to
\cite{assaf_composite_2005}.

@@

\subsection{Correction methods}

Three out of four methods presented in \autoref{sec:intro}
were tested on the evaluation framework.

\Gls*{fmb} correction
is essentially the same as the one used for generating the 
distortion, using the synthetic fieldmap to infer a new \gls*{vsm}.

@@

\subsection{Evaluation}

The four modules described so far (namely, phantom
deformation, \gls*{fmb} correction, \gls*{reb}
correction and \gls*{t2b} correction) were conveniently
integrated in second order workflows using
\emph{nipype} \cite{gorgolewski_nipype:_2011}.
The choice of this tool grants the reproducibility
of the experiments, among other advantages.

To evaluate the influence of
distortions on the final connectivity outcome,
an intermediate step was also implemented to
produce deterministic tractography,
using the \emph{Diffusion Toolkit} \cite{wang_diffusion_2007}.
The original phantom, one distorted version, and
the corrected instances are fed to this module.
Additionally, the original tissue probability maps
are also distorted and corrected, for computing overlap
indices.
Each tractography was restricted to its corresponding 
\gls*{wm} mask, after distortion and correction when
applied. For all the cases, we used 10 random seeds at each seeding
voxel (belonging to one out of 27 different seeding regions).
The seeding regions were taken from the 
ground-truth, mimicking the usual procedure on real data
(where regions are typically mapped from the
anatomically correct T1).

>>>>

\subsection{Digital Phantom}
Based on the fiber geometries of the digital phantom 
created for the High angular resolution diffusion imaging 
reconstruction Challenge held in ISBI 2013 
(San Francisco, US), we simulated high resolution 
(0.5mm isotropic) T1 (TE/TR=10/1500ms) and T2 
(TE/TR=90/5000ms) images, as well as a two \gls*{dmri}
images (1.0mm isotropic, $b$=1200, 1 B0 image) with
32 and 64 evenly-distributed directions.
Diffusion is modeled by a restricted and a hindered
compartment, similar to \cite{assaf_composite_2005}.
The phantom includes \gls*{wm} fiber bundles,
\gls*{gm} and \gls*{csf}. Physical properties
(T1/T2 times in ms) used in simulation are
($0.832\pm0.010$ / $79\times10^{-3}\pm0.6\times10^{-3}$)
for \gls*{wm},
($1.331\pm0.013$ / $110\times10^{-3}\pm2.0\times10^{-3}$) for
\gls*{gm} and ($3.5\pm0.1$ / $0.25\pm0.01$) for \gls*{csf}.

@@

\subsection{Correction methods}
\label{sec:correction}
Three out of four methods presented in \autoref{sec:intro}
were tested on the evaluation framework.

\Gls*{fmb} correction
is essentially the same as the one used for generating the 
distortion, on the inverse direction. Normally distributed 
noise was added to the fieldmap (for a signal-to-noise 
ratio of 20dB) before correction.

@@

\subsection{Evaluation}
\label{sec:evaluation}
The original phantom, one distorted version, and
the corrected instances are then connected to a
\gls*{dwi} reconstruction and tractography pipeline.
Additionally, the original tissue probability maps
are also distorted and corrected to provide tractography
with the required \gls*{wm} masks.
These maps are also used in a final assessment module.

The framework provides \gls*{dwi} reconstruction and
deterministic tractography,
with \emph{Diffusion Toolkit} \cite{wang_diffusion_2007}
for diffusion tensor imaging (configured with 10 random
seeds per voxel by default), and \emph{MRTrix}
\cite{tournier_mrtrix:_2012} for high angular resolution
diffusion tensor imaging (default parameters set to use
constrained spherical deconvolution, maximum harmonic
order of 6, 150000 desired tracks).
For both options, the seeding regions can be set to use
the distorted-corrected \gls*{wm} mask or the regions used
to generate the ground-truth, this second case mimicking
the usual procedure on real data (where regions are 
typically mapped from the anatomically correct T1).

### 2013-10-10 Publications/2013-ISBI14/04-results.tex
Updating for second deadline

\section{RESULTS}
\label{sec:results}
The results of the proposed experiments are summarized
in \autoref{fig:results} and \autoref{table:results}.
The remaining of this section provides extended descriptions
and interpretation of results.

@@

\subsection{Geometrical correctness}

\subsection{Signal loss recovery}

A second workflow investigated the similarity of the recovered
signal with respect to the original (undistorted).
In this second study, \gls*{reb} performed
better than the other two methods, as reported
in \autoref{table:results}. \gls*{reb} scored a $91.0\%$
similarity index for the \textit{b0} volume and an average $97.65\%$
for the remaining 31 \glspl*{dwi}. Again, the second
qualified was \gls*{fmb}, which achieved very close results
for the \glspl*{dwi} ($96.26\%$) but not as good for the \textit{b0}
volume. Visual inspection of the recovered data confirms the 
results presented in the Table (\autoref{fig:results}, first
row). Finally, \gls*{t2b} confirmed that its correction
can not be compared, even when using the intensity compensation
based on the determinant of the deformation field.

@@

@@

Even though all the surveyed methods produced visually 
sound results, our study suggested that \gls*{reb} is the 
most accurate methodology for correction of susceptibility-induced
distortions in \gls*{dwi}. Moreover, \gls*{reb} also achieved
the best results in terms of connectivity, being able to recover
one of two lost connections after distortion. \Gls*{fmb} achieved 
very positive scores. Finally, the \gls*{t2b} method did not achieve
the necessary high-standards to recommend its use. We understand
that specific methods with anisotropic regularization that completely
restrict deformations to the phase-encoding direction would perform
significantly better compared to the standard method included
in this study.

>>>>

\section{RESULTS}
\label{sec:results}

All the modules described in the previous Section were
conveniently integrated in workflows using \emph{nipype}
\cite{gorgolewski_nipype:_2011}. The choice of this tool
grants the reproducibility of the experiments,
and the evaluation workflow is publicly released.
The results of the proposed experiments are summarized
in \autoref{fig:results} and \autoref{table:results}.
The remaining of this section provides extended descriptions
and interpretation of results.

\subsection{Geometrical correctness and signal recovery}

@@

A second workflow investigated the similarity of the recovered
signal with respect to the original (undistorted).
In this second study, \gls*{reb} performed
better than the other two methods, as reported
in \autoref{table:results}. \gls*{reb} scored a $91.0\%$
similarity index for the \textit{b0} volume and an average $97.65\%$
for the remaining 31 \glspl*{dwi}. Again, the second
qualified was \gls*{fmb}, which achieved very close results
for the \glspl*{dwi} ($96.26\%$) but not as good for the \textit{b0}
volume. Visual inspection of the recovered data confirms the 
presented numerical results (\autoref{fig:results}, first row).

@@

Results derived from diffusion tensor imaging 
(using 32 directions \glspl*{dwi})
were completely hindered by the high complexity of the
fibers contained in the phantom (crossing, fanning, etc.).
The immediate conclusion is that, with highly complex anatomies,
connectivity outcome is more sensitive to 
reconstruction and tractography than distortions. Therefore,
all results presented are based on the 64-directions dataset
reconstructed with constrained spherical deconvolution.

@@

Even though all the surveyed methods produced visually 
sound results, our study suggested that \gls*{reb} is the 
most accurate methodology for correction of susceptibility-induced
distortions in \gls*{dwi}. Moreover, \gls*{reb} also achieved
the best results in terms of connectivity, being able to recover
one of two lost connections after distortion. \Gls*{fmb} achieved 
very positive scores. Finally, the \gls*{t2b} method did not achieve
the necessary high-standards to recommend its use. We understand
that specific methods with anisotropic regularization that completely
restrict deformations to the phase-encoding direction would perform
significantly better compared to the standard method included
in this study.
%Finally, \gls*{t2b} confirmed that its correction
%can not be compared, even when using the intensity compensation
%based on the determinant of the deformation field.
